### 2014-11-03 Implementations.tex
Updated Sensor Data Privacy Impact.
Added dashboard implementation description

\chapter{Implementation of Privacy Protection Measures}
\subsection{Privacy Policy}

@@

\textbf{What Personal Information do we Collect?}\\
The Live+Gov Service Center stores the following personal information:\\
- Name\\
- Email Address\\

The Mobile Sensor Collection Application collects the following data from your mobile phone:\\
- Sensor Data from Accelerometer (50Hz)\\
- Sensor Data from GPS sensors (~every 30 sec)\\
- Daily Human Activities (walking, standing, sitting, etc.) (every 30sec)\\
- Used Service Lines (bus, train, tram) (every 2 minutes)\\

@@

@@

\section{Processing Journal}
* name of the controller
* purpose of processing
* description of the data categories
* recipients of the data if disclosed
* transfer to third countries
* description of security of processing

The journal is available for anybody to access via Email.

Reporting to Data Protection Authority (Bundesamt für Datenschutz)?

\section{Privacy Dashboard}
TODO: Add description of dashboard. With relation to 7C's of Privacy.

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/login.png}
\caption{Live+Gov Privacy Dashboard Login}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/recordings.png}
\caption{Live+Gov Privacy Dashboard Recording Overview}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/raw.png}
\caption{Live+Gov Privacy Dashboard Raw Data View}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/processed.png}
\caption{Live+Gov Privacy Dashboard Processed Data View}
\end{figure}

\section{Transfer Encryption (HTTPS/SSL)}
\begin{figure}
\includegraphics[width=\textwidth]{screenshots/HTTPS.png}
\caption{HTTPS connection to Privacy Dashboard}
\end{figure}

\section{Database Access Policies}
Access rules for data controller and processor.
Data Processing Journal.
Python script for deletion of data when it is older than a given date.

\section{Anonymization of trips}
Privacy by Design: We do never collect name and email address in our DB.
Python script that deletes user information from trip.
% Python script that reomves precise date information, only day of week.

>>>>

\chapter{Implementation of Privacy Protection Measures}
\section{Privacy Policy}

@@

\textbf{What Personal Information do we Collect?}\\
The Mobile Sensor Collection Application collects the following data from your mobile phone:\\
- Sensor Data from Accelerometer (50Hz)\\
- Sensor Data from GPS sensors (~every 30 sec)\\
- Daily Human Activities (walking, standing, sitting, etc.) (every 30sec)\\
- Used Service Lines (bus, train, tram) (every 2 minutes)\\

@@

The Live+Gov Services does NOT store or collect names and email addresses.

@@

\section{Privacy Dashboard}
The Live+Gov Privacy Dashboard is a software component that allows the
citizens to view the data about themselves that is stored in the
Sensor Data Storage server.  The foundation of the Privacy Dashboard
component is the inspection tool that was introduced in D1.1 as a
means to perform basic quality controlling of the collected data.

The Privacy Dashboard is currently endowed with the following views:
\begin{itemize}
  \item \textbf{Login Screen} (cf. Figure \ref{fig:PDLogin}).  Before
    access to the dashboard is granted, the user is prompted for
    credentials.  The credentials are generated by the sensor
    collection application on the mobile device.  The username can be
    freely chosen by the user. The password is then generated by the
    mobile application and transferred to the server alongside with
    the recording.
 \item \textbf{Recording Overview.} (Cf. Figure \ref{fig:PDOverview}).
   This view shows all recordings that the user submitted to the Data
   Storage Service including basic meta information like start data of
   recording, duration. The user can click on each recording in order
   to view the recorded data in the Raw Data View.

   An important feature of the recording overview table is the delete
   button (cross on the very left.) That allows the user to
   selectively delete recordings, that shall not be analyzed by the
   Live+Gov system. Internally the click on the delete button only
   flags the corresponding trip for deletion. The actual deletion is
   carried out as a batch job on a regular bases. This addition step
   prevents accidental loss of data.

 \item \textbf{Raw Data View.} (cf. Figure \ref{fig:PDRawData}) The
   raw data view shows all recorded raw data to the user.  GPS data is
   visualized on a map. Motion sensor data is displayed as plots. Wifi
   and GSM data is displayed in lists.
   The raw data from each sensor can be downloaded individually as CSV file.

 \item \textbf{Processing View.} (cf. Figure
   \ref{fig:PDProcessedData}).  The processing view shows the data
   mining end-products for each processing step.  The Figure shows the
   processing view for the Activity Recognition. On the top a
   time-line is show that visualizes the different recognized
   activities via color-coding. A map below displays the corresponding
   GPS track with the same color coding.

   A similar view for Service Line Detection results is currently
   under development.
\end{itemize}

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/login.png}
\caption{Live+Gov Privacy Dashboard Login}
\label{fig:PDLogin}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/recordings.png}
\caption{Live+Gov Privacy Dashboard Recording Overview}
\label{fig:PDOverview}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/raw.png}
\caption{Live+Gov Privacy Dashboard Raw Data View}
\label{fig:PDRawData}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{screenshots/processed.png}
\caption{Live+Gov Privacy Dashboard Processed Data View}
\label{fig:PDProcessedData}
\end{figure}

\section{Transfer Encryption (HTTPS/SSL)}
The data transfer from the mobile device to the Data Storage component
as well as the transfer to the Privacy Dashboard relay on HTTP. In
order to avoid eavesdropping by externals that have access to the
communication channels, we secured all HTTP connections with SSL
encryption.

The implementation of this measure was rather straight forward. A SSL
certificate was issued by the University of Koblenz, and installed to
our webservers. Requests to this servers can now be done using HTTPS.
Figure \ref{fig:PDHTTPS} shows a screenshot of such a secured
connection to our Privacy Dashboard. The necessary adjustments to the
Sensor Collection Component were minimal, since Android comes with
native support for HTTPS connections.

\begin{figure}[h]
\includegraphics[width=\textwidth]{screenshots/HTTPS.png}
\caption{HTTPS connection to Privacy Dashboard}
\label{fig:PDHTTPS}
\end{figure}

\section{Data Expiring Dates}
Trips stored in the Sensor Data Storage Service have been endowed with
a data expiration date. A python script is provided that performs the
of expired data. In our installation we have this script running as a
cron job every night. The resulting schema of the trip table listed in
Figure \ref{fig:TripScheme}.

By default the expiration data is set to one year after the recording
was uploaded. We are working on including a feature for setting the
expiration date for each recording inside the privacy dashboard.

\begin{figure}
\centering
\begin{verbatim}
  Column  |          Type
----------+------------------------
 trip_id  | integer
 user_id  | character varying(36)
 start_ts | bigint
 stop_ts  | bigint
 name     | character varying(255)
 deleted  | boolean
 expires  | bigint
\end{verbatim}
\caption{Database Schema of the Trip Table}
\label{fig:TripScheme}
\end{figure}

\section{Anonymization of trips}
The Sensor Data Storage Service follows a privacy by design guideline
in that it does not store personal information like names and email
address at all.

In the case that such information should be linked to the user-id
stored in the trip table we provide a python script that randomizes
the user-ids and thereby removes all links that might have existed
before.

\begin{figure}[fontsize=\tiny]
\centering
\begin{verbatim}
Usage: privacy.py [-d database name] [-u database user]
                  [-p database password]
                  [-h database host] [-o operation]

Valid Operations:

cleanup: Flags all expired rows as deleted
    No Options

anonymization: Sets all user_ids to 0
    Option:
        SQL Select Query: Every row found by this query will
        be set to user_id = 0.

haircut: Cuts trips where they split and could be traceable
    Options:
        First - the radius in which we look
        Second - the number of users that needs to be in the
                 radius

blur: Offsets every GPS point by a random amount
    Options:
        The maximum distance a point can have to its origin

k_anonymity: Creates a centroid for every GPS point
    Options:
        First - Number of maximum GPS points to use. If 0 then
                we grab every point inside the radius
        Second - Maximum distance a point can have to be used
                for the creation of the new point
\end{verbatim}
\caption{Usage options for the privacy.py script.}
\end{figure}
